\documentclass{article}

\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{multicol}
\usepackage[skip=3pt]{caption}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{subcaption}
\usepackage{fancyhdr}
\usepackage{tabulary}
\usepackage{times}
\usepackage{datetime}
\usepackage{titling}
%\pagestyle{fancy}
\usepackage[most]{tcolorbox}
\usepackage[T1]{fontenc}
\usepackage[font=small,labelfont=bf,tableposition=top]{caption}
\DeclareCaptionLabelFormat{andtable}{#1~#2  \&  \tablename~\thetable}

\usepackage{algorithm}
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

\newcommand{\tab}[1][1]{\noindent \hspace{#1cm} }

\usepackage[style=numeric]{biblatex}
\addbibresource{myreferences.bib}

\begin{document}

\noindent \textbf{CSCE 586 - Design and Analysis of Algorithms}

\noindent \textbf{Date:}  \today 

\noindent \textbf{Name:}  Micah Hayden

\noindent \textbf{Assignment:}  Final Exam - Take Home Portion, Version A

\noindent \textbf{Documentation:} 

\hrulefill

\section{Hidden Surface Removal Problem:}
\subsection{Which algorithmic paradigm will you use to solve this problem?}  
I will utilize Divide-and-Conquer to solve this problem.

\subsection{Why did you chose the algorithmic paradigm selected above to solve this problem?}
I believe this paradigm is suited to divide and conquer.  Each line has similar characteristics, following the same equation $y_i = a_i \cdot x + b_i$.  The problem can be divided into small, independent sub-problems using the slope of each line.  The base case occurs when $n \leq 3$:  because no 3 lines intersect at a single point, the resulting "uppermost" lines can be found in constant time.
 
\subsection{Give an algorithm that takes $n$ lines as input, and in $O(n \log{n})$ time returns all of the lines that are visible.  Provide a clear description of the algorithm.}

Let $L$ be a set of lines, $|L|=n$, where $L_i=m_i \cdot x + b_i$.

\begin{algorithm}
\caption{Hidden-Surface-Removal:  HSR($L$)}
\begin{algorithmic}
\State Sort $L$ by ascending slope, such that $L_i$ has slope $m_i$ and $m_i < m_{i+1}$ for all $i$.
\State $|L| = n$
\If {$n \leq 2$}
	\State \Return The set of lines $L$, and their intersection $a$.
\ElsIf {$n = 3$}
	\State Let $a =$ the intersection of $L_1$ and $L_3$
	\State Let $b =$ the intersection of $L_1$ and $L_2$
	\If {$x_b < x_a$}
		\State Let $c =$ the intersection of $L_2$ and $L_3$.
		\State \Return $L$ and $\{b, c\}$
	\Else 
		\State \Return $L - \{L_2\}$ and $\{a\}$
	\EndIf
\Else
	\State $L, A = HSR(\{L_1, \dots , L_{\frac{n}{2}}\})$
	\State $L', B = HSR(\{L_{\frac{n}{2}+1}, \dots , L_n\})$
\EndIf
\State Merge $A$ and $B$ into $C$ by increasing $x$ coordinate
\State Find the first element in $C$ for which $L' > L$, denote this element as $c_k$.
\State Let $L_i \in L$ be the uppermost line in $L$ immediately before $c_k$.
\State Let $L_j \in L'$ be the uppermost line in $L'$ immediately after $c_k$.
\State Let point $p_{int}$ be the intersection of lines $L_i$ and $L_j$.
\State $L = \{L_1, L_2, \dots, L_i\} \cup \{L'_j, L'_{j+1}, \dots L_n \}$
\State $C = \{A_1, \dots A_{i-1} \} \cup p_{int} \cup \{B_j, \dots B_{n-1} \}$
\State \Return $L$ and $C$
\end{algorithmic}
\end{algorithm}

\subsection{Perform asymptotic analysis of your algorithm's running time.  Also, consider the run time performance of a best case, worst case, and average case input model scenario.}

\noindent \textbf{Worst Case:}

\noindent \textbf{Best Case:}

\noindent \textbf{Average Case:}

\subsection{Provide a proof that your algorithm works correctly:}

\newpage

\section{Bipartite Matching Problem}

\subsection{Which algorithmic paradigm best describes this algorithm?}  

\subsection{Why did you choose the algorithmic paradigm selected above?}

\subsection{Give an example of a bipartite graph $G$ for which this algorithm does not return the maximum matching.}

\subsection{Let $M$ and $M'$ be matchings in a bipartite graph $G$.  Suppose that $|M'| > 2 \cdot |M|.$  Show that there is an edge $e' \in M'$ such that $M \cup {e'}$ is a matching in $G$.}

\subsection{Using the previous claim (and your supporting proof) to further prove that the algorithm is optimal or that the algorithm is $\rho$-optimal approximate (in this case be sure to derive the value of $\rho$ as part of your proof).}


\newpage

\section{Number Partitioning Problem}

\subsection{Problem Statement:}  
Show that the \textit{Number Partitioning} is NP-complete using the \textit{Subset Sum} problem.

\subsection*{Show \textit{Number Partitioning} $\in$ NP}
Let $Y = Number \, Partitioning$

\subsection*{Choose an NP-Complete problem $X$:}
Let \textit{Subset Sum} be X.

\subsection*{Prove that $X \leq_P Y$}


\printbibliography

\end{document}
